Differential geometric algebra with Leibniz and Grassmann

### 2019-11-01 paper/paper.tex
revised Project.toml, Makie.jl links, and revised paper grammar

\maketitle
\begin{figure}[ht]
%\centerline{\includegraphics[width=4cm]{grassmann.png}}
\centerline{\includegraphics[width=4cm]{../docs/src/assets/logo.png}}
\caption*{github.com/chakravala/Grassmann.jl}
\end{figure}
\begin{abstract}
	The \textit{Grassmann.jl}
	package provides tools for computations based on multi-linear algebra and spin groups using the extended geometric algebra known as Leibniz-Grassmann-Clifford-Hestenes algebra.
	Combinatorial products include
	exterior, regressive, inner, and geometric; along with the Hodge star, adjoint, reversal, and boundary operators.
	The kernelized operations are built up from composite sparse tensor products and Hodge duality, with high dimensional support for up to 62 indices using staged caching and precompilation. Code generation enables concise with highly extensible definitions.
	\textit{DirectSum.jl}
	multivector parametric type polymorphism is based on tangent vector spaces and conformal projective geometry.
	Additionally, the universal interoperability between different sub-algebras is enabled by \textit{AbstractTensors.jl},
	on which the type system is built.
	%\headingtable
\end{abstract}
\begin{itemize}
	\item \textbf{DirectSum.jl}: Abstract tangent bundle vector space types (unions, intersections, sums, etc.)
	\item \textbf{AbstractTensors.jl}: Tensor algebra abstract type interoperability with vector bundle parameter
	\item \textbf{Grassmann.jl}: $\langle$Leibniz+Grassmann-Clifford-Hestenes$\rangle$ differential geometric algebra of multivector forms
	\item \textbf{Leibniz.jl}: Derivation operator algebras for tensor fields
	\item \textbf{Reduce.jl}: Symbolic parser generator for Julia expressions using REDUCE algebra term rewriter
\end{itemize}
Mathematical foundations and some of the nuances in the definitions specific to the \textit{Grassmann.jl} implementation are concisely described, along with the accompanying support packages that provide an extensible platform for computing with geometric algebra at high dimensions.
The design is based on the \verb`TensorAlgebra` abstract type interoperability from \textit{AbstractTensors.jl} with a \verb`VectorBundle` type parameter from \textit{DirectSum.jl}. Abstract vector space type operations happen at compile-time, resulting in a differential conformal geometric algebra of hyper-dual multivector forms.

@@

\begin{definition}
	Define differential $d:\Omega^p V\ra\Omega^{p+1}V$ and co-differential $\delta:\Omega^pV\ra\Omega^{p-1}V$ such that \cite{bishop-goldberg}
	\begin{align*}
		\star d\omega &= \star(\nabla\wedge\omega) = \nabla\times\omega, & \omega\cdot\nabla &= \omega\vee\star\nabla = \partial\omega =-\delta\omega.
	\end{align*}
	These two maps have the special properties $d\circ d=0$ and $\partial\circ\partial = 0$ for any form $\omega$ and vector field $\nabla$. 
	In topology there is \textit{boundary} operator $\partial$ defined by $\partial\epsilon = \epsilon\cdot\nabla = \sum_k\partial_k\epsilon_k$ and is commonly discussed in terms the limit $\epsilon(x)\cdot\nabla\omega(x) = \lim_{h\ra0} \frac{\omega(x+h\epsilon)-\omega(x)}{h}$, which is the directional derivative \cite{sobczyk}.
\end{definition}
\begin{example}
	[Vorticity of vector-field]
	$\star d(dx_1+dx_2+dx_3) = (∂_2 -∂_3)dx_1 + (∂_3 -∂_1)dx_2 + (∂_1 -∂_2)dx_3$.
\end{example}
\begin{example}
	[Boundary of 3-simplex]
	Faces of simplex (oriented): $\partial(w_{1234}) = -\partial_4w_{123}+\partial_3w_{124}-\partial_2w_{134}+\partial_1w_{234}$.
\end{example}

@@

For the null-basis, complement operations are different:
\begin{align*}
	\star v_\infty &= \star(v_++v_-) = (v_- + v_+)v_{1...n} = v_{\infty1...n} \\
 	\star 2v_\emptyset &= \star(v_--v_+) = (v_+ - v_-)v_{1...n} = -2v_{\emptyset1...n}
\end{align*}
The Hodge complement satisfies $\gen{\omega\ast\omega}I=\omega\wedge\star\omega$. This property is naturally a result of using the geometric product in the definition.
An additional metric independent version of the complement operation is available with the \verb`!` operator,
\begin{align*}
	!v_\infty &= !(v_++v_-) = (v_- - v_+)v_{1...n} = 2v_{\emptyset1...n} \\
 	!2v_\emptyset &= !(v_--v_+) = (v_+ + v_-)v_{1...n} = -v_{\infty1...n}
\end{align*}
That variation of complement satisfies $\lambda I = \omega\,\wedge\,!\omega$ instead.
\begin{example}%[\textit{using Leibniz $\bigoplus\,$Grassmann}]
	S"$\infty\emptyset\text{+++"}(\nabla$\verb`^`$2) \,\,  \mapsto\, \, -2\partial_{\infty\emptyset} + \partial_1^2 + \partial_2^2 + \partial_3^2 $
\end{example}
Let $\nabla\in\Lambda^1V$, then $\omega = (\nabla\bs\nabla)\ominus\omega = \nabla\bs(d\omega + \partial\omega)$ where $\nabla\parallel\partial\omega$ and $\nabla\perp d\omega$.
Let's reflect across the hyperplane $\star\nabla$,
\begin{align*}
	\nabla\bs (d\omega-\partial\omega) &= \nabla\bs(d\omega-\partial\omega)\ominus(\nabla\bs\nabla) \\
									   &= -\nabla^2\bs(d\omega+\partial\omega)\ominus\nabla = -\nabla\bs\omega\ominus\nabla. %= \omega\oslash\nabla
\end{align*}
Hence, reflection by hyperplane $\star\nabla$ has isometry $\omega\oslash\nabla$ which for $\nabla = v_j$ is the map $\mb R^n \ra \mb R_1\times\cdots\times\conj{\mb R}_j\cdots\times\mb R_n$.

@@

\textbf{Acknowledgements}: Many thanks for discussions with Steven De Keninck, Utensil Song, Alexander Arsenovic, Hugo Hadfield, Karl Wessel; the helpful Julia community (e.g. Jameson Nash, Stefan Karpinski, Jeff Bezanson, Kristoffer Carlsson, Petr Krysl, Nathan Smith, etc); and thanks for support from my family.

>>>>

\maketitle
\begin{figure}[ht]
\centerline{\includegraphics[width=4cm]{img/grassmann.png}}
%\centerline{\includegraphics[width=4cm]{../docs/src/assets/logo.png}}
\caption*{github.com/chakravala/Grassmann.jl}
\end{figure}
\begin{abstract}
	The \textit{Grassmann.jl}
	package provides tools for computations based on multi-linear algebra and spin groups using the extended geometric algebra known as Leibniz-Grassmann-Clifford-Hestenes algebra.
	Combinatorial products include
	exterior, regressive, inner, and geometric; along with the Hodge star, adjoint, reversal, and boundary operators.
	The kernelized operations are built up from composite sparse tensor products and Hodge duality, with high dimensional support for up to 62 indices using staged caching and precompilation. Code generation enables concise yet highly extensible definitions.
	\textit{DirectSum.jl}
	multivector parametric type polymorphism is based on tangent vector spaces and conformal projective geometry.
	Additionally, the universal interoperability between different sub-algebras is enabled by \textit{AbstractTensors.jl},
	on which the type system is built.
	%\headingtable
\end{abstract}
\begin{itemize}
	\item \textbf{DirectSum.jl}: Abstract tangent bundle vector space types (unions, intersections, sums, etc.)
	\item \textbf{AbstractTensors.jl}: Tensor algebra abstract type interoperability with vector bundle parameter
	\item \textbf{Grassmann.jl}: $\langle$Leibniz+Grassmann-Clifford-Hestenes$\rangle$ differential geometric algebra of multivector forms
	\item \textbf{Leibniz.jl}: Derivation operator algebras for tensor fields
	\item \textbf{Reduce.jl}: Symbolic parser generator for Julia expressions using REDUCE algebra term rewriter
\end{itemize}
Mathematical foundations and some of the nuances in the definitions specific to the \textit{Grassmann.jl} implementation are concisely described, along with the accompanying support packages that provide an extensible platform for computing with geometric algebra at high dimensions.
The design is based on the \verb`TensorAlgebra` abstract type interoperability from \textit{AbstractTensors.jl} with a \verb`VectorBundle` type parameter from \textit{DirectSum.jl}. Abstract vector space type operations happen at compile-time, resulting in a differential conformal geometric algebra of hyper-dual multivector forms.

@@

\begin{definition}
	Define differential $d:\Omega^p V\ra\Omega^{p+1}V$ and co-differential $\delta:\Omega^pV\ra\Omega^{p-1}V$ such that \cite{bishop-goldberg}
	\begin{align*}
		\star d\omega &= \star(\nabla\wedge\omega) = \nabla\times\omega, & \omega\cdot\nabla &= \omega\vee\star\nabla = \partial\omega =-\delta\omega.
	\end{align*}
	These two maps have the special properties $d\circ d=0$ and $\partial\circ\partial = 0$ for any form $\omega$ and vector field $\nabla$. 
	In topology there is \textit{boundary} operator $\partial$ defined by $\partial\epsilon = \epsilon\cdot\nabla = \sum_k\partial_k\epsilon_k$ and is commonly discussed in terms the limit $\epsilon(x)\cdot\nabla\omega(x) = \lim_{h\ra0} \frac{\omega(x+h\epsilon)-\omega(x)}{h}$, which is the directional derivative \cite{sobczyk}.
\end{definition}
\begin{example}
	[Vorticity curl of vector-field]
	$\star d(dx_1+dx_2+dx_3) = (∂_2 -∂_3)dx_1 + (∂_3 -∂_1)dx_2 + (∂_1 -∂_2)dx_3$.
\end{example}
\begin{example}
	[Boundary of 3-simplex]
	Faces of simplex (oriented): $\partial(w_{1234}) = -\partial_4w_{123}+\partial_3w_{124}-\partial_2w_{134}+\partial_1w_{234}$.
\end{example}

@@

For the null-basis, complement operations are different:
\begin{align*}
	\star v_\infty &= \star(v_++v_-) = (v_- + v_+)v_{1...n} = v_{\infty1...n} \\
 	\star 2v_\emptyset &= \star(v_--v_+) = (v_+ - v_-)v_{1...n} = -2v_{\emptyset1...n}
\end{align*}
The Hodge complement satisfies $\gen{\omega\ast\omega}I=\omega\wedge\star\omega$. This property is naturally a result of using the geometric product in the definition.
An additional metric independent version of the complement operation is available with the \verb`!` operator,
\begin{align*}
	!v_\infty &= !(v_++v_-) = (v_- - v_+)v_{1...n} = 2v_{\emptyset1...n} \\
 	!2v_\emptyset &= !(v_--v_+) = (v_+ + v_-)v_{1...n} = -v_{\infty1...n}
\end{align*}
For that variation of complement, $||\omega||^2 I = \omega\,\wedge\,!\omega$ holds.
\begin{example}%[\textit{using Leibniz $\bigoplus\,$Grassmann}]
	S"$\infty\emptyset\text{+++"}(\nabla$\verb`^`$2) \,\,  \mapsto\, \, -2\partial_{\infty\emptyset} + \partial_1^2 + \partial_2^2 + \partial_3^2 $
\end{example}
Let $\nabla\in\Lambda^1V$, then $\omega = (\nabla\bs\nabla)\ominus\omega = \nabla\bs(d\omega + \partial\omega)$ where $\nabla\parallel\partial\omega$ and $\nabla\perp d\omega$.
Let's reflect across the hyperplane $\star\nabla$,
\begin{align*}
	\nabla\bs (d\omega-\partial\omega) &= \nabla\bs(d\omega-\partial\omega)\ominus(\nabla\bs\nabla) \\
									   &= -\nabla^2\bs(d\omega+\partial\omega)\ominus\nabla = -\nabla\bs\omega\ominus\nabla. %= \omega\oslash\nabla
\end{align*}
Hence, reflection by hyperplane $\star\nabla$ has isometry $\omega\oslash\nabla$ which for $\nabla = v_j$ is the map $\mb R^n \ra \mb R_1\times\cdots\times\conj{\mb R}_j\cdots\times\mb R_n$.

@@

\textbf{Acknowledgements}: Many thanks for discussions with Steven De Keninck, Utensil Song, Alexander Arsenovic, Hugo Hadfield, Karl Wessel; the helpful Julia community (e.g. Jameson Nash, Stefan Karpinski, Jeff Bezanson, Kristoffer Carlsson, Petr Krysl, Nathan Smith, Hongying Li, etc); and thanks for support from my family.
